---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# DGBD

<!-- badges: start -->
<!-- badges: end -->

Discrete Generalized Beta Distributions (DGBDs) are an alternative option to model
Rank-Abundance Distributions in ecology. Power-law distributions are accurate 
for an intermediate range. Yet, they don't fit the tail ends where the behavior
is logarithmic. While they can be modified by additional parameters, 
DGBDs better model the tail regions of the distribution.


DGBDs not only fit well to Rank-Abundance Distributions, but also to ordering 
distributions in other biological systems and fields. Read further in Martinez-Mekler et al. (2009) https://doi.org/10.1371/journal.pone.0004791
DGBDs are described by 3 parameters: 'A', 'a' and 'b'. This 'R' package
fits a DGBD to a given set of abundance data, reports these parameters and
uses 'ggplot2' to graph the model.


## Installation

You can install the development version of DGBD like so:

``` r
devtools::install_github("Fa-Moe/DGBD")
```

## Usage

You can quickly analyze philoseq-class objects and data frames.

```{r example}
library(DGBD)
Tara_Data
Tara_DGBD <- BC_multiple(Tara_Data,is_phyloseq=TRUE,columns_select_multiple = c(1,3),plot_silent=T)

BC_report(hmp_wgs,2)
```
